Testing Charts: viewer's perceptual accuracy in surveys

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
update to figure and t-test

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: 1               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Heike
    first-init: H.
    last: Hofmann
    marker: 2
  - first: Nola
    first-init: N.
    last: duToit
    marker: 1
  - first: Edward
    first-init: E.
    last: Mulrow
    marker: 1
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: NORC at the University of Chicago
    country: United States
    marker: 1
  - institution: Iowa State University
    department: Statistics
    country: United States
    marker: 2
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
  \usepackage{booktabs}
  \usepackage{graphicx}
  \usepackage{hyperref}
  \usepackage{setspace}
documentclass: jds
format:
  html: default
  pdf: 
    include-in-header: authors.tex
    keep-tex: true
editor: 
  markdown: 
    wrap: sentence
---

@@

\kr{Can we update the figure and t-test with combined data across all tasks/all rounds 1-3?}

```{r combine_rounds_1_and_2}

lambda <- calculate_lambda(round1_adj$weight, round2_adj$weight)
round_12 <- round1_adj %>%
            mutate(weight = lambda[1]*weight) %>%
            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
    rbind(round2_adj %>%
            mutate(weight = (1-lambda[1])*weight) %>%
            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
round_12_long <- round_12 %>%
  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")

survey_12 <-
  survey::svydesign(
  data = round_12_long,
  ids = ~CaseId, weights = ~weight)
```

```{r ttest}
paired <- svyttest(Response~Question, survey_12)
abcd <- svyby(~Response, by=~Question, survey_12, svytotal)

@@

\kr{ Can we update this to include the values for ROund 3 as well?}
Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis.
Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16).

\kr{We need to update this chart to include the data from all 3 rounds instead of just the first 2}

@@

gg1 <- round_12_long %>%
  mutate(
    Response = ifelse(Response, "Correct", "Not correct"),
    Response = factor(Response, levels=c("Not correct", "Correct")),
    Question = ifelse(Question=="CorrectAB",
                      "Unaligned: A or B?", "Aligned: C or D?")
  ) %>%
#  filter(Response == "Correct") %>%
  ggplot(aes(x = Question, fill = Response, weight = weight)) +
  geom_bar() +
  ylab("Number of (effective) Respondents") +
  xlab("") +
  scale_fill_manual(values=c("grey", "forestgreen")) +
  scale_y_continuous(
      sec.axis = sec_axis(~./sum(round_12$weight)*100,
                          name="Percent responses")
  ) +
  theme(legend.position = "bottom") +
  ggtitle("(a) Accuracy of survey responses")  +
  geom_rect(
    aes(
      x = rev(c("Aligned: C or D?", "Unaligned: A or B?")),
      xmin = 2:1 - 0.45,
      ymin = ResponseTRUE - 1.96* se.ResponseTRUE,
      xmax = 2:1 + 0.45,
      ymax = ResponseTRUE + 1.96* se.ResponseTRUE,
      weight = 1),
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = marginal_results
  )

gg2 <- round_12 %>%
  mutate(
    `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"),
    `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct")
    ) %>%
  pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>%
  pcp_scale() %>%
  pcp_arrange() %>%
  ggplot(aes_pcp()) +
    geom_pcp_axes() +
    geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+
    geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) +
    geom_pcp_labels(colour = "black") +
  scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25),
                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
                     sec.axis = sec_axis(~.*100,
                          name="Percent responses")) +
  xlab("") +
  scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) +
  ggtitle("(b) Combinations of responses") +
  theme(legend.position = "none")

>>>>

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: 1               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Heike
    first-init: H.
    last: Hofmann
    marker: 2
  - first: Nola
    first-init: N.
    last: duToit
    marker: 1
  - first: Edward
    first-init: E.
    last: Mulrow
    marker: 1
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: NORC at the University of Chicago
    country: United States
    marker: 1
  - institution: Iowa State University
    department: Statistics
    country: United States
    marker: 2
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
  \usepackage{booktabs}
  \usepackage{graphicx}
  \usepackage{hyperref}
  \usepackage{setspace}
documentclass: jds
format:
  pdf: 
    include-in-header: authors.tex
    keep-tex: true
  html: default
editor: 
  markdown: 
    wrap: sentence
---

@@

\kr{Can we update the figure and t-test with combined data across all tasks/all rounds 1-3?}\hh{done}

```{r ttest}
paired <- svyttest(correct~task, survey_123)
abcd <- svyby(~correct, by=~task, survey_123, svytotal)

@@

\kr{ Can we update this to include the values for ROund 3 as well?}\hh{done}
<!--Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambdas[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.-->

Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis.
Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16). \hh{how about: $p$-value: $<< 0.0001$ -- the e-16 looks very computery}

\kr{We need to update this chart to include the data from all 3 rounds instead of just the first 2} \hh{done}

@@

gg1 <- round_123_long %>%
  mutate(
    Response = ifelse(correct, "Correct", "Not correct"),
    Response = factor(Response, levels=c("Not correct", "Correct")),
    Question = ifelse(task=="ab",
                      "Unaligned: A or B?", "Aligned: C or D?")
  ) %>%
#  filter(Response == "Correct") %>%
  ggplot(aes(x = Question, fill = Response, weight = weight)) +
  geom_bar() +
  ylab("Number of (effective) Respondents") +
  xlab("") +
  scale_fill_manual(values=c("grey", "forestgreen")) +
  scale_y_continuous(
      sec.axis = sec_axis(~./sum(round_123$weight)*100,
                          name="Percent responses")
  ) +
  theme(legend.position = "bottom") +
  ggtitle("(a) Accuracy of survey responses")  +
  geom_rect(
    aes(
      x = rev(c("Aligned: C or D?", "Unaligned: A or B?")),
      xmin = 2:1 - 0.45,
      ymin = correctTRUE - 1.96* se.correctTRUE,
      xmax = 2:1 + 0.45,
      ymax = correctTRUE + 1.96* se.correctTRUE,
      weight = 1),
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = marginal_results
  )

gg2 <- round_123 %>%
  mutate(
    `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"),
    `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct")
    ) %>%
  pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>%
  pcp_scale() %>%
  pcp_arrange() %>%
  ggplot(aes_pcp()) +
    geom_pcp_axes() +
    geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+
    geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) +
    geom_pcp_labels(colour = "black") +
  scale_y_continuous("Number of (effective) Respondents", breaks=seq(0,1,by=0.25),
                     labels = round(sum(round_123$weight)*seq(0,1,by=0.25),0),
                     sec.axis = sec_axis(~.*100,
                          name="Percent responses")) +
  xlab("") +
  scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) +
  ggtitle("(b) Combinations of responses") +
  theme(legend.position = "none")
